%  Voorwoord (dankwoord) en toelating tot bruikleen
\addcontentsline{toc}{chapter}{Preface}
\newpage

\noindent \textbf{\huge Preface}

\vspace{1.5cm}

\noindent

This work was completed with the help of several people, I would like to thank. First of all my supervisors, prof. dr. ir. Joni Dambre and my counsellor ir. Thibault Verhoeven for their guidance and patience during the year. They gave me the freedom to steer the direction of my thesis, while still ensuring that the end result has scientific value. Additionally, I would like to thank my external supervisor, dr. ir. Pieter Van Mierlo for evaluating my work. I would also like express appreciation for the ideas and suggestions that my thesis colleges, Nina, Bjorn and Thijs for gave me during this work.
\npar
This preface wouldn't be complete without thanking my parents, stepparents, brothers and sister for giving me their support, not only this year, but throughout all my years as a student. I would also like to thank my mother and father for giving me the opportunity to study and standing by me.
\npar
Last but not least, I would also like to thank my friends, among others, Cedric, Patrick, Michiel, Thomas and Wouter for supporting me over the years of my study and creating a pleasant working atmosphere. 

\addvspace{3cm}

\hfill \noindent Andreas De Lille

\hfill \noindent Ghent, 1 June 2016

\clearpage

\noindent \textbf{\huge Persmission for Usage}
\addcontentsline{toc}{chapter}{Permission for Usage}
\vspace{1.5cm}

\noindent
``The author(s) gives (give) permission to make this master dissertation available for consultation
and to copy parts of this master dissertation for personal use.
 In the case of any other use, the copyright terms have to be respected, in particular with regard to
the obligation to state expressly the source when quoting results from this master dissertation.``
\npar
``De auteur geeft de toelating deze scriptie voor consultatie beschikbaar
te stellen en delen van de scriptie te kopi\"eren voor persoonlijk
gebruik.\\
Elk ander gebruik valt onder de beperkingen van het auteursrecht,
in het bijzonder met betrekking tot de verplichting de bron uitdrukkelijk
te vermelden bij het aanhalen van resultaten uit deze scriptie.''

\addvspace{4cm}

\hfill \noindent Andreas De Lille

\hfill \noindent Ghent, 1 June 2016

%\addvspace{12cm}
%\doclicenseThis